---
title: "High Frequency Trading Data - High Fidelity Milestone"
author: "Viktor Ludwig, Markus Tretzmüller, Michael Trimmel"
date: "29 11 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r, echo=FALSE, message=FALSE}
library(orderbook)
library(Rserve)

#Rserve()


actualPath <- getwd()
pathToData <- paste(actualPath, "/Desktop/Visualisation/", sep = "")

messageFile <- "AAPL_2012-06-21_34200000_57600000_message_5.csv"
orderbookFile <- "AAPL_2012-06-21_34200000_57600000_orderbook_5.csv"




######### load  data ########################
# load  order  book  data
dataOB<-read.csv( paste(pathToData, orderbookFile, sep = ""))
# load  message  data
dataM <- read.csv( paste(pathToData, messageFile, sep = ""))
# merge two  datasets
data <- cbind(dataM , dataOB[ ,-1])
####### load  completed ###################



nlevels <- (dim(dataOB)[2] - 1) / 4
colnms <- c("Time", "Type", "OrderID", "Size", "Price", "TradeDirection")

for(i in 1:nlevels)
{
    colnms <- c(colnms, paste("ASKp", i, sep = ""), paste("ASKv", i, sep = ""), paste("BIDp", i, sep = ""), paste("BIDv", i, sep = ""))
    
}
colnames(data) <- colnms

data[1:5,]
plot(data[,'Time'], data[,'Price'], type = "l", lwd=1, ylab = "price", xlab = "time")
```


## Team 

- Viktor Ludwig
- Markus Tretzmüller
- Michael Trimmel


## Impementation Details

We start exploring our data with a mixture of R and Tableau, exploring our data as easily as possible. For more flexibility in the final visualization we then will change to D3. 

## Milestones

+15.11.2016 First Teammeeting with Torsney-Weir Thomas for fist informations what to do.

+15.11.2016 - 24.11.2016 Getting information about tradig.

+24.11.2016 Meeting with Goekhan Cebiroglu, and the VIS Team to get spezific information.

+24.11.2016 - 15.12.2016 LoFi Prototype and HiFi Prototype in Cran and Tableau

+15.12.2016 M3 HiFi Presentation 13h00

+15.11.2016-24.01.2017 Finalize Visualization

## Who did what 

+ Briefing the team members in order book dynamics (Viktor Ludwig)
+ Getting deeper knowledge of order books. (Viktor Ludwig, Markus Tretzmüller, Michael Trimmel)
+ Website (Markus Tretzmüller)
+ Dashboards: Each of us designed one dashboard. Pro and cons were discussed in group.
+ Report: Michael Trimmel


## Dashboard 1 (Viktor Ludwig)

\includegraphics[width=480pt]{Dashboard_Viktor.jpg}

### Discussion

What we want to show with this dashboard. A overview over the price changes over the day like in the chart at the bottom. In this chart we can chance the resolution of the other views with independent sliders. A pro for that kind of slider is that someone can change the time of the day that he wants to inspect and additionaly the time resolution. A con of the independent slider is that there has to be precomputed a really large amount of data. 

At the right upper side of the dashboard we want to show the state of the order book with a bar chart, which is seperated in a "bid" and "ask" region. A negative viewpoint of that chart is that the labels of the x-Axes is changing here for every timestep because the price of the best bid and asks is always in motion.

At the right bottom we show the ratio of all "bid" and "ask" orders. In our view, the negative point of that view is that the visualization in tachometer style resembles a quantity in physics or engineering and might be confusing. The pro of that view is that you can easily figure out the amount of positive or negative ratio.


## Dashboard 2 (Markus Tretzmüller)

\includegraphics[width=480pt]{Dashboard_Markus.png}

### Discussion

What we want to show with this dashboard. The chart on the bottom depicts the actual price over time. This view implements brush & and filter functionality and sets the other views according to the range selected. The central view is the most enhanced and most complex view. 

It shows the price distribution of placed orders over time due a box plot. Additionally the volumes of placed orders is depicted at the bottom of these views. The third view depicts the type of orders executed (whether submission, cancellation or execution) with a pie chart. 

## Dashboard 3 (Michael Trimmel)

\includegraphics[width=480pt]{Dashboard_Mike.jpg}

### Discussion

What we want to show with this dashboard? We want to show the order flow of stocks. This is achieved with multiple views, showing different metrics of the market flow. The main window shows the actual price of the stock which is common in visualizations of stocks and trading software. Here, the user can change the the resolution of the price and the time with multiple sliders in each direction.
This is a big advantage, but, the computational costs for that will be huge to filter that out. Maybe a few basic timescales which can be chosen by checking the box at the upper right corner can be precomputed and shown directly when the user is selecting it to solve the problem.

The buy sell order ratio shows the actual trend of the order flow. This is shown like a histogram chart, directly in main window which saves space of the overall dashboard, but, if the ratio is very high and the price of the stock is low there can be conflicts between the 2 charts in one window.



The descriptiv statistics box at the right bottom is very useful to get specific information about the selected "bid" or "ask" price. I want to show more detailed information for users who wants to get the exact price or amount of a selected price level.

The checkbox at the top middle is rejected. A first idea was that the user can compare 2 different stocks



## References

### Websites 

+ Lobster: https://lobsterdata.com/
+ Forum: http://money.stackexchange.com

### Papers and Presentations

+ https://www.princeton.edu/~rcarmona/download/short_courses/Princeton_June2013/RTG1.pdf
+ http://vda.univie.ac.at/Teaching/Vis/15s/final_reports/team_3.pdf
+ https://lobsterdata.com/info/docs/LobsterReport.pdf

### Software

+ http://things-about-r.tumblr.com/
+ http://www.tableau.com/
+ https://cran.r-project.org/
+ https://d3js.org/


https://cran.r-project.org/web/packages/obAnalytics/vignettes/guide.html
https://cran.r-project.org/web/packages/orderbook/vignettes/orderbook.pdf


## Issues 

Basically understanding of economics and the mechanics of stock market. Getting familiar with the idea of high frequency trading.

## Further steps

It would be cool to predict the future trend. It would also be nice to create a edge bundled chart for the transaction flows, but the issue is that the trader don´t have to show his name or MPID.
